### 2019-03-13 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Again lots of work on Chapter 6, see #14

A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} \cite{Auer02NonStochastic}. Under this model, rewards are completely arbitrary and are not assumed to follow any probability distribution.
For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied. The pseudo regret of the EXP3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}. However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round. For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
It is sometimes referred to as the abruptly changing MAB, or switching environments \cite{MellorShapiro13}.
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $i$ at round $t$ has some mean $\mu_i(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $i_t^* = \arg\max_{i} \mu_i(t)$.

In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 \cite{Auer02} used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

@@

A policy $\pi$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$:
%
\begin{equation}\label{def:6:nti}
    R_T^{\pi} = \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
\end{equation}

In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint by
\begin{equation}
    \tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}.
\end{equation}
%
Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arms are \iid{} (hence the name of \emph{piece-wise stationary} problems).

@@

Introducing the number of change-points on arm $i$, defined as $\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$, it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.

@@

\TODOL{Speak about the article \cite{WeiSrivastava18} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}

@@

\paragraph{Advantages of our approach.}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
This study is supported by both theoretical and empirical results presented in this chapter.
%
In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{i,t} |\mu_i(t) - \mu_i(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{i,j} |\mu_i(t) - \mu_j(t)|$.

@@

@@

The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl}, we can show that in any one-dimensional exponential family, we have
\[\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right],\]
%
where for $k \leq k'$, $\widehat{\mu}_{k:k'} = \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
%
This motivates the following definition of the Bernoulli GLRT change point detector.

@@

Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR} but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence between two elements in that exponential family that have mean $x$ and $y$. For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to \eqref{def:6:GLR} with $\kl(x,y) = 2(x-y)^2$ when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.

In this section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.
Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $2(x-y)^2$.

% ----------------------------------------------------------------------------
\subsection{Non-Asymptotic Properties of the B-GLRT}\label{subsec:6:PropGLR}

@@

\TODOL{We need to PROVE our results for false alarm and delay !.}

@@

In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small. To define $\beta$, we need to introduce the function $\cT$, defined for $x>0$ by
%
\begin{equation}\label{def:6:function_T}
    \cT(x) ~:=~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
\end{equation}
where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$.
%
And for any $x \ge 0$, $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise. The function $\cT$ is easy to compute numerically.
The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96} (inverse of $x \mapsto x \exp(x)$), as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.
%
Its use for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on the function $\cT$ are also given:  $\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.
% The proof of Lemma~\ref{lem:6:FalseAlarm}, that actually holds for any sub-Bernoulli distribution, is given below.

@@

Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
\begin{equation}
    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
\end{equation}
where $\phi\mu(\lambda)$ is the log moment generating of some one-dimensional exponential family. The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).

Let's go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}.
A first key observation is
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.

@@

@@

\begin{proof}
    \TODOL{Write it!}
\end{proof}

@@

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{equation}
    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
\end{equation}
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{}-like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
the number of selections of arm $i$, and
\[\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t) = \ln(t) + 3 \ln(\ln(t))$ (if $t>1$ else $f(t)=0$),
the index is defined as
%
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \bigr\}.
\end{equation}

>>>>

A first possibility to cope with non-stationary is to model the decision making problem as an \emph{adversarial bandit problem} \cite{Auer02NonStochastic}. Under this model, rewards are completely arbitrary and are not assumed to follow any probability distribution.
For adversarial environments, the pseudo-regret, which compares the accumulated reward of a given strategy with that of the best fixed-arm policy, is often studied. The pseudo regret of the EXP3 algorithm has been shown to be $\cO(\sqrt{KT})$, which matches the lower bound given by \cite{Auer02NonStochastic}. However, this model is a bit too general for the considered applications, where reward distributions do not necessarily vary at every round. For these reasons, an intermediate model, called the \emph{piece-wise stationary MAB}, has been introduced by \cite{Kocsis06} and \cite{YuMannor09}.
It is sometimes referred to as the abruptly changing \cite{WeiSrivastava18Abruptly}, or switching environments \cite{MellorShapiro13}.
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $i$ at round $t$ has some mean $\mu_i(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $i_t^* = \arg\max_{i} \mu_i(t)$.

In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

@@

A policy $\pi$ chooses the next arm to play based on the sequence of past plays and obtained rewards.
The performance of $\pi$ is measured by its (piece-wise stationary) \emph{regret}, the difference between the expected reward obtained by an oracle policy playing an optimal arm $i^*_t$ at time $t$, and that of the policy $\pi$,
%
\begin{equation}\label{def:6:nti}
    R_T^{\pi} = \E\left[\sum_{t=1}^T \left(\mu_{i^*_t}(t) - \mu_{I_t}(t)\right)\right].
\end{equation}

In the piece-wise \iid{} model, we furthermore assume that there is a (relatively small) number of \emph{breakpoints}, denoted by $\Upsilon_T:=\sum_{t=1}^{T-1} \indic\left(\exists i\in\{1,\dots,K\}: \mu_t(i) \neq \mu_{t+1}(i)\right)$.
If we denote $\tau^{(0)} = 0$, we define the $k$-th breakpoint recursively by
\begin{equation}
    \tau^{(k)} = \inf\{t > \tau^{(k-1)} : \exists i : \mu_i(t) \neq \mu_{i}(t+1)\}.
\end{equation}
%
Hence for $t\in[\tau^{(k)} + 1,\tau^{(k+1)}]$, that is on any stationary segments, the rewards $(X_{i,t})$ associated to each arm are \iid{} (hence the name of \emph{piece-wise stationary} problems).

@@

We define the \emph{number of change-points} on arm $i$,
$\NCi := \sum_{t=1}^{T-1} \indic\left(\mu_t(i) \neq \mu_{t+1}(i)\right)$,
for which it clearly holds that
$\NCi \leq \Upsilon_T$, but there can be an arbitrary difference between these two quantities for some arms. Letting $C_T := \sum_{i=1}^K \NCi$ be the total number of change-points on the arms, one can have $C_T \in \{ \Upsilon_T, \dots, K\Upsilon_T \}$.
We illustrate the two extreme cases in problems $1$ and $2$ presented in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} below.

@@

@@

\paragraph{Advantages of our approach.}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
% This study is supported by both theoretical and empirical results presented in this chapter.
%
In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{i,t} |\mu_i(t) - \mu_i(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{i,j} |\mu_i(t) - \mu_j(t)|$.

@@

\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
%
In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
we will assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
%
This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.

% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}

@@

The Generalized Likelihood Ratio statistic for this test is defined by
\[\GLR(n) :=\frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \ldots, X_n ; \mu_0)},\]
%
where $\ell(X_1, \ldots, X_n ; \mu_0)$ and $\ell(X_1, \ldots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the likelihoods of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$ over $\cH_1$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$).
%
\begin{equation}\label{eq:6:BernoulliDivergence}
    \kl(x,y) := x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right).
\end{equation}
%
Indeed, as proved in Appendix~\ref{app:6:GLR_with_kl}, we can show that in any one-dimensional exponential family, we have
\[\GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right],\]
%
where for $k \leq k'$, $\widehat{\mu}_{k:k'} := \frac{1}{k - k' + 1} \sum_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.
%
This motivates the following definition of the Bernoulli GLRT change point detector.

@@

Asymptotic properties of the GLRT for change-point detection have been studied by \cite{LaiXing10} for Bernoulli distributions and more generally for one-parameter exponential families, for which the GLR test is defined as in~\eqref{def:6:GLR}, but with $\kl(x,y)$ replaced by the Kullback-Leibler divergence $d(x,y)$ between two elements in that exponential family that have mean $x$ and $y$.
For example, the Gaussian GLR studied by \cite{Maillard2018GLR} corresponds to the stopping time \eqref{def:6:GLR} with the choice $d(x,y) = 2(x-y)^2$, when the variance is set to $\sigma^2=1/4$, and non-asymptotic properties of this test are given for any $(1/4)$-subGaussian samples.
Note that Pinsker's inequality gives that $\kl(x,y) \geq 2(x-y)^2$, hence the B-GLRT may stop earlier that the Gaussian GLR, based on the quadratic divergence $d(x,y) = 2(x-y)^2$.
%
In the next section, we provide new non-asymptotic results about the B-GLRT under the assumption that the samples $(X_t)$ come from a sub-Bernoulli distribution, which holds for any distribution supported in $[0,1]$.

% ----------------------------------------------------------------------------
\subsection{Non-asymptotic properties of the B-GLRT}\label{subsec:6:PropGLR}

@@

@@

In Lemma~\ref{lem:6:FalseAlarm} below, we propose a choice of the threshold function $\beta(n,\delta)$ under which the probability that there exists a \emph{false alarm} under \iid{} data is small. To define $\beta$, we need to introduce the function $\cT$, defined for $x>0$ by
%
\begin{equation}\label{def:6:function_T}
    \cT(x) ~:=~ 2 \widetilde{h}\left(\frac{h^{-1}(1+x) + \ln(2\zeta(2))}{2}\right),
\end{equation}
where for $u \ge 1$ we define $h(u) = u - \ln(u)$ and its inverse $h^{-1}(u)$,
%
we define $\widetilde{h}(x) = e^{1/h^{-1}(x)} h^{-1}(x)$ if $x \ge h^{-1}(1/\ln (3/2))$ and $\tilde{h}(x) = (3/2) (x-\ln(\ln (3/2)))$ otherwise, for any $x \ge 0$,
and with the value $\zeta(2) = \pi^2 / 6$.
%
Even if it does not have a closed form expression, the function $\cT$ is easy to compute numerically.
%
The inverse $h^{-1}$ can be computed using $\mathcal{W}$, the Lambert $\mathcal{W}$ function \cite{Corless96}, which is the inverse of $x \mapsto x \exp(x)$, as $h^{-1}(x) = - \mathcal{W}(- \exp(-x))$.

The use of $\cT$ for the construction of concentration inequalities that are uniform in time is detailed in~\cite{KK18Martingales}, where tight upper bound on this function $\cT$ are also given:
$\cT (x) \simeq x + 4 \ln\big(1 + x + \sqrt{2x}\big)$ for $x\geq 5$ and $\cT(x) \sim x$ when $x$ is large.

@@

\begin{proof}
Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
\begin{equation}
    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
\end{equation}
where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family. The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).

Let us go back to the Bernoulli case with divergence given in \eqref{eq:6:BernoulliDivergence}.
We first observe that
% A first key observation is
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\mu \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)
\end{align*}%
\begin{align*}
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$:
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6\ln(1+\ln(s+r))+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3\ln(1+\ln(s))+3\ln(1+\ln(r)) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
%
\end{proof}

@@

This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.

@@

\begin{proof}
    \TODOL{Write it ?}
\end{proof}

@@

The \GLRklUCB{} algorithm can be viewed as a \klUCB{} algorithm allowing for some \emph{restarts} on the different arms. A restart happens when the B-GLRT change-point detector detects a change on the arm that has been played (line $9$).
%
To be fully specific, $\GLR_\delta(Z_1,\dots,Z_n) = \mathrm{True}$ if and only if
\begin{equation}
    \sup_{1 < s < n} \left[s \times \kl \left(\frac{1}{s}\sum_{i=1}^s Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right) + (n-s) \times \kl \left(\frac{1}{n-s}\sum_{i=s+1}^t Z_i , \frac{1}{n}\sum_{i=1}^n Z_i\right)\right] \geq \beta(n,\delta),
\end{equation}
%
with $\beta(n,\delta)$ defined in \eqref{def:6:beta}, or $\beta(n,\delta) = \ln(3n^{3/2}/\delta)$, as recommended in practice, see Section~\ref{sub:6:PracticalConsiderations}. We define the (\klUCB{} like) index used by our algorithm, by denoting
$\tau_i(t)$ the last restart that happened for arm $i$ before time $t$,
\[n_i(t) = \sum_{s=\tau_i(t)+1}^{t} \indic(A_s=i)\]
the number of selections of arm $i$, and
\[\widehat{\mu}_i(t) = \frac{1}{n_i(t)} \sum_{s=\tau_i(t)+1}^{t}X_{i,s} \indic(A_s=i)\]
their empirical mean (if $n_i(t)\neq0$).
%
With the exploration function $f(t)$ defined as $\ln(t) + 3 \ln(\ln(t))$ if $t>1$ and $f(t)=0$ otherwise,
the index is defined as
%
\begin{equation}\label{def:6:UCB}
    \UCB_i(t) := \max \Bigl\{ q\in[0,1] : n_i(t) \times \kl\left(\widehat{\mu}_i(t),q\right) \leq f(t - \tau_i(t)) \Bigr\}.
\end{equation}

%-----------------------------------------------------------------------------
\paragraph{Two options for restarts.}
%
For this algorithm, we simultaneously investigate two possible behavior: \emph{global restart} (reset the history of all arms once a change was detected on one of them, line $11$), and \emph{local restart} (reset only the history of the arm on which a change was detected, line $13$), which are the two different options in Algorithm~\ref{algo:6:GLRklUCB}.
%
Under local restart, in the general case the times $\tau_i(t)$ are not equal for all arms, hence the index policy associated to \eqref{def:6:UCB} is \emph{not} a standard \UCB{} algorithm, as each index uses a \emph{different exploration rate}.
%
One can highlight that in the \CUSUMUCB{} algorithm, which is the only existing algorithm based on local restart, the \UCB{} index are defined differently\footnote{~This alternative choice is currently not fully supported by theory, as we found mistakes in the analysis of \CUSUMUCB: Hoeffding's inequality is used with a \emph{random} number of observations and a \emph{random} threshold to obtain Eq. $(31)$-$(32)$.}:
$f(t-\tau_i(t))$ is replaced by $f(n_t)$ with $n_t = \sum_{i=1}^K n_i(t)$.

%-----------------------------------------------------------------------------
\paragraph{Threshold function $\beta$.}
